---
title: "README"
author: "Genevieve Jacobs"
date: "2/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Homework: Use the bridges data to find features that you think would be predictive of unemployment. Fit a linear model to predict the number of unemployed. Fit another model to predict the unemployment rate. Then, use the unemployed number and rate from the previous month as additional predictors. How do things change?

```{r}

rm(list =ls())

## Load the libraries and data

library(dplyr)
library(tidyverse)
library(ggplot2)
library(blscrapeR)

## Download all states' bridges data

read_csv("2018HwyBridgesDelimitedAllStates.txt")
blsbridgesData = read_csv("2018HwyBridgesDelimitedAllStates.txt")
bls_december = get_bls_county()

```


```{r}

## Make the variables and data frames we want:

filtered_states = blsbridgesData %>% filter(STATE_CODE_001 <= "12") 
  ## This is actually the first 10, not 12, states of the US alphabetically. Refer to the STATE CODE section of bridges data manual.

joined = filtered_states %>% left_join(bls_december, by = c("COUNTY_CODE_003" = "fips_county", "STATE_CODE_001" = "fips_state")) %>%  select(labor_force, SUFFICIENCY_RATING, COUNTY_CODE_003, YEAR_BUILT_027, STATE_CODE_001, unemployed, unemployed_rate, ROADWAY_WIDTH_MT_051, ADT_029)

filtered_frame = as.data.frame(joined)

```



```{r}

## Unemployment Count Linear Model:

unemploymentcount_model = lm(unemployed ~ SUFFICIENCY_RATING + ROADWAY_WIDTH_MT_051 + ADT_029, data = filtered_frame)

summary(unemploymentcount_model)

```


```{r}
## Unemployment Rate Linear Model:

unemployment_rate_model = lm(unemployed_rate ~ SUFFICIENCY_RATING + ROADWAY_WIDTH_MT_051 + ADT_029, data = filtered_frame)

summary(unemployment_rate_model)

```

 
```{r}
## Include November Data in Model: 

bls_november = get_bls_county("November 2018")

joined_november = filtered_states %>% left_join(bls_november, by = c("COUNTY_CODE_003" = "fips_county", "STATE_CODE_001" = "fips_state")) %>%  select(labor_force, SUFFICIENCY_RATING, COUNTY_CODE_003, YEAR_BUILT_027, STATE_CODE_001, unemployed, unemployed_rate, ROADWAY_WIDTH_MT_051, ADT_029)

unemploymentcount_model_total = lm(unemployed ~ SUFFICIENCY_RATING + ROADWAY_WIDTH_MT_051 + ADT_029 + joined_november$unemployed, data = filtered_frame)

summary(unemploymentcount_model_total)
```

WRITE-UP:

The variables in the bridges data are mostly categorical. I wanted to include quantitative variables in my linear model so that I could better interpret the numerical values and their meaning. I also wanted to choose quantitative variables so that I would not have to interpret upwards of 9 dummy variables. The quantitative variables I predicted would be best for predicting unemployment rates and unemployment counts were sufficiency rating, roadway width, and average daily traffic. These variable names are somewhat self-explanatory. The sufficiency rating is the quality of each bridge updated yearly. I thought sufficiency rating would have an indirect correlation with unemployment rating/count: if the bridge is doing well, then it is probably because workers have been upkeeping it. The roadway width is simply how wide the bridge is. I theorized that a wider bridge means that more people can use the bridge to get to a job or otherwise seek opportunities. The average daily traffic is a quantifier for the estimate of traffic a bridge has in one day. I chose this thinking that more daily traffic means that more people are using the bridge to go to their job. I believed each of these three variables would be indirectly correlated with unemployment.

For the unemployment rate model, these variables were all significant. Sufficiency rating and average daily traffic were negatively correlated with unemployment rate. This was as I hypothesized. However, roadway width had a positive beta value. Roadway width turned out to have a positive beta value for all of the models. I think this is reasonable. Cities will have wider bridges, and cities tend to have higher rates of unemployment, so this positive beta for roadway width is not too out there.

For the unemployment count model, all variables were significant and had positive beta values. This simply goes to show that count is not a good measurement for a variable like unemployment. High sufficiency rating, high average daily traffic, a wide bridge all correlate with higher populations. Cities get more updates (higher sufficiency), more traffic (higher average daily traffic), and need wider bridges to support their population (roadway width). Cities have higher populations, thus, more unemployed. All in all, it is best to use rates to interpret unemployment results.

When I included November unemployment data in the linear model, I found that the R-squared value skyrocketed from about 11% to >95%. This makes sense. Unemployment rates will not change drastically month-to-month, so of course this model can explain a lot of the variation. Likewise, the beta value for the november data was positive. I was surprised to find that all of the prior variables were still significant. Once again, roadway width was positive, but sufficiency rating and average daily traffic were negative. 







